---
title: "Practice4_DA5030"
author: "Shivam Verma"
output: pdf_document
---

```{r setup, include=TRUE}
library(tm)
library(SnowballC)
library(wordcloud)
library(e1071)
library(gmodels)
```

Problem - 1

Step-2 Exploring and Preparing the data
```{r}
# reading the dataset
sms_raw <- read.csv("spammsg.csv", stringsAsFactors = FALSE)
# exploring characteristics of data
str(sms_raw)
sms_raw$type <- factor(sms_raw$type)
table(sms_raw$type)
```

Data Preperation - cleaning and standardizing text data
```{r}
# creating a volatile corpus
sms_corpus <- Corpus(VectorSource(sms_raw$text))
#printing the result
print(sms_corpus)
#summarize specific messages
inspect(sms_corpus[1:3])
# standardizing the messages to lowercase characters
corpus_clean <- tm_map(sms_corpus, tolower)
# removing all the numbers from the corpus
corpus_clean <- tm_map (corpus_clean , removeNumbers)
# removing the stopwords
corpus_clean <- tm_map(corpus_clean, removeWords, stopwords())
# removing the punctuation
corpus_clean <- tm_map(corpus_clean, removePunctuation)
# removing the white spaces
corpus_clean <- tm_map(corpus_clean, stripWhitespace)
# stemming variants of words
corpus_clean <- tm_map(corpus_clean, stemDocument)
inspect(corpus_clean[1:3])
# creating DTM by applying tokenization
sms_dtm <- DocumentTermMatrix(corpus_clean)
```

Data Preperation - creating trianing and test datasets
```{r}
# dividing the data into two portions: 75 percent for training and 25 percent for testing
sms_raw_train <- sms_raw[1:4180,]
sms_raw_test <- sms_raw[4181:5574,]
# creating test and train labels
sms_train_label <- sms_raw[1:4180,]$type 
sms_test_label <- sms_raw[4181:5574,]$type
sms_dtm_train <- sms_dtm[1:4180,]
sms_dtm_test <- sms_dtm[4181:5574,]
sms_corpus_train <- corpus_clean[1:4180]
sms_corpus_test <- corpus_clean[4181:5574]
prop.table(table(sms_raw_train$type))
prop.table(table(sms_raw_test$type))
# Visualizing text data - word clouds
wordcloud(corpus_clean, min.freq = 50, random.order = FALSE)
# making subsets of raw data by the type
spam <- subset(sms_raw, type=="spam")
ham <- subset(sms_raw, type=="ham")
# creating word cloud for each subset
wordcloud(spam$text, max.words = 40, scale= c(3,0.5))
wordcloud(ham$text, max.words = 40, scale = c(3,0.5))
```

Data preperation - creating indicator features for frequent words 
```{r}
# eliminating words that appear in less than 5 SMS messages 
sms_freq_words <- findFreqTerms(sms_dtm_train, 5)
str(sms_freq_words)
# filtering DTM test and train data
sms_dtm_freq_train<- sms_dtm_train[ , sms_freq_words]
sms_dtm_freq_test <- sms_dtm_test[ , sms_freq_words]
# function to convert counts to Yes/No strings
convert_counts <- function(x)
{
  x <- ifelse(x>0, "Yes", "No")
}
# Applying the function to all columns
sms_train <- apply(sms_dtm_freq_train, MARGIN = 2, convert_counts)
sms_test <- apply(sms_dtm_freq_test, MARGIN = 2, convert_counts)

```

Step 3 - training a model on the data
```{r}
# Applying Naive Bayes algorithm 
sms_classifier <- naiveBayes(sms_train, sms_train_label) 
# Making predictions
sms_test_pred <- predict(sms_classifier, sms_test)
```

Step - 4 Evaluating model performance
```{r}
# generating cross table for performance evaluation
CrossTable(sms_test_pred, sms_test_label, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))

```

We can observe only 29 SMS predicted wrong. thus accuracy is approx 98%.

Step 5 - improving the model performance
```{r}
# trying to improve performance by setting Laplace estimator
sms_classifier2 <- naiveBayes(sms_train, sms_train_label, laplace = 1) 
sms_test_pred2 <- predict(sms_classifier2, sms_test)
CrossTable(sms_test_pred, sms_test_label, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))

```
It is obsereved the performance remained same even after introducing the Laplace estimator.

Problem - 2

```{r}
library(klaR)
# loading the iris dataset of R
data(iris)
# finding the number of rows in iris data
nrow(iris)
# Getting the summary of data
summary(iris)
# Creating a view of data
head(iris)
# This statement acts as a loop from 1 to length of dataset to fetch numbers which are divisible by 5.
testidx <- which(1:length(iris[, 1]) %% 5 == 0)

# creating train dataset excluding those rows which are included in test dataset
iristrain <- iris[-testidx,]
# creating test dataset with the previously fetched numbers divisible by 5
iristest <- iris[testidx,]

# applying Naive Bayes Algorithm
nbmodel <- NaiveBayes(Species~., data=iristrain)

# checking for the accuracy
# Using the test data to make prediction for species of test data
prediction <- predict(nbmodel, iristest[,-5])
table(prediction$class, iristest[,5])
```
It is observed that there are only two misclassifications which gives accuracy of about 93%
