---
title: "Practice3_DA5030"
author: "Shivam Verma"
output: pdf_document
---

```{r setup, include=FALSE}
library(caret)
library(class)
library(gmodels)
```
Implementing KNN using Class package.


Step- 2 Preparing and exploring the data
```{r}
c <- read.csv("cancer.csv", stringsAsFactors = FALSE)
# Checking if data is structured
str(c)
# Removing ID form data set as it don't provide useful information
c <- c[-1]
head(c)
# Getting count distribution of pateints
table(c$diagnosis_result)
# Renaming Variables
c$diagnosis <- factor(c$diagnosis_result, levels = c("B", "M"), labels = c("Benign", "Malignant"))
# Returning results as percentage
round(prop.table(table(c$diagnosis)) * 100, digits = 1)
```
Normalizing the data set.
```{r}
set.seed(1234)
normalize <- function(x) {
  return ((x - min(x)) / (max(x) - min(x))) }
c_n <- as.data.frame(lapply(c[2:9], normalize))
summary(c_n$radius)
```
Creating training and test data set.
```{r}
c_train <- c_n[1:65,]
c_test <- c_n[66:100,]
# Now considering daignosis factor into consideration.
c_train_labels <- c[1:65, 10]
c_test_labels <- c[66:100, 10]
```
Step-3 Training a model on data
```{r}
# Considering k as the square root of the number of observations.
c_test_pred <- knn(train = c_train, test = c_test, cl = c_train_labels, k=10)
```
Step 4 – Evaluating the model performance
```{r}
CrossTable(x= c_test_labels, y= c_test_pred, prop.chisq = FALSE )
```
Out of 35 cases 7 are True Positive, 15 are True Negative (when positive class is Benign). Accuracy is determined by (TP+TN)/Total cases i.e. approx. 63%, thus there is room for improvement. 
```{r}
#Confusion Matrix for class:knn
confusionMatrix(c_test_labels, c_test_pred)
```
Implementing KNN using Caret package 

Using the same data splitted into 65:35 as train and test, as used before.

```{r}
trctrl <- trainControl(method = "repeatedcv", number = 10, repeats = 3)
set.seed(1234)
# Data Scaling Performed (Standardization using z-scores)
knn_fit <- train(c_train_labels ~., data = cbind(c_train_labels, c_train), method = "knn",
 trControl=trctrl,
 preProcess = c("center", "scale"),
 tuneLength = 10)
#printing the model
knn_fit
#Predicted values
c_caret_pred <- predict(knn_fit, newdata = c_test)

```

```{r}
CrossTable(x= c_test_labels, y= c_caret_pred, prop.chisq = FALSE)
```
Out of 35 cases 10 are True Positive, 16 True Negative (when positive class is Benign). Accuracy is determined by (TP+TN)/Total cases i.e. approx. 75%.
```{r}

plot(knn_fit)

```

```{r}
confusionMatrix(c_test_labels, c_caret_pred)
```
















