---
title: "Practice 2_DA5030"
author: "Shivam Verma"
output: pdf_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(tidyverse)

```
Question 1.
Determine which states are outliers in terms of murders. Outliers, for the sake of this question, are defined as values that are more than 1.5 standard deviations from the mean.
```{r}
data <- USArrests

# z-score calculation
data$z_score <- abs(((mean(data$Murder)) - data$Murder) / (sd(data$Murder)))

# filter outliers
outliers <- data[which(data$z_score > 1.50), 0:4 ]
outliers
```
Question 2. 
Is there a correlation between urban population and murder, i.e., as one goes up, does the other statistic as well? Comment on the strength of the correlation. Calculate the Pearson coefficient of correlation in R.
```{r}
cor.test(x = data$Murder, y = data$UrbanPop, method = "pearson")
```
As value for pearson correlation estimates to be around 0.07 this shows a weak linear relationship between urban population and murder. As one increases, the other typically doesn't increase. 


Question 3. 
Forecast phone use for the next time period using a 2-year weighted moving average (with weights of 5 for the most recent year, and 2 for other), exponential smoothing (alpha of 0.4), and linear regression trendline.
```{r}
# Weighted Avg Forecast
df<- read.csv("mobile.csv")
n<- nrow(df)
last2 <- df[n:(n-1), 2]
w <- c(5,2)
sw <- w*last2
Ft_wavg <- round(sum(sw)/sum(w))
Ft_wavg

# Exponential Smoothing Forecast
a <- 0.4
df$Ft <- 0
df$E <- 0

df$Ft[1] <- df$Subscribers[1]
for(i in 2:n)
{
  df$Ft[i] <- df$Ft[i-1] + a * df$E[i-1]
  df$E[i] = df$Subscribers[i] - df$Ft[i]
   
}
Ft_esp <- round(df$Ft[n] + a * df$E[n])
Ft_esp

# Liner Regression Forecast
ggplot(df, aes(x=Year, y = Subscribers)) + geom_line() 
model <- lm(df$Subscribers ~ df$Year)
summary(model)
print(model)
Ft_lm <- -15710760 + 18276748 *(n+1)
Ft_lm
```
Question 4.
Calculate the squared error for each model, i.e., use the model to calculate a forecast for each given time period and then the squared error. Finally, calculate the average (mean) squared error for each model. Which model has the smallest mean squared error (MSE)?
```{r}
# Weighted Avg
df<- read.csv("mobile.csv")
df$Ft <- 0
df$SqErr <- 0
df$Ft[1] <- df$Subscribers[1]
df$Ft[2] <- df$Subscribers[2]
w <- c(5,2)
for(i in 3:n)
{
  last2 <- df[(i-1): (i-2), 2]
  sw <- w*last2
  df$Ft[i]<- (sum(sw)/sum(w))
  df$SqErr[i] <- (df$Subscribers[i] - df$Ft[i])^2
}
MSE_wavg <- mean(df$SqErr)
MSE_wavg


# Exponential Smoothing model
df<- read.csv("mobile.csv")
a <- 0.4
df$Ft <- 0
df$E <- 0

df$Ft[1] <- df$Subscribers[1]
for(i in 2:n)
{
  df$Ft[i] <- df$Ft[i-1] + a * df$E[i-1]
  df$E[i] = df$Subscribers[i] - df$Ft[i]
   
}
df$SqErr <- 0
for(i in 1:n)
{
  df$SqErr[i] <- (df$E[i])^2
  
}
MSE_esp <- mean(df$SqErr)
MSE_esp 


#Linear Regression model
df<- read.csv("mobile.csv")
df$Ft <- 0
df$SqErr <- 0
for(i in 1:n)
{
  df$Ft[i] <- (-15710760) + 18276748 *(i)
  df$SqErr[i] <- (df$Subscribers[i] - df$Ft[i])^2
  
}
MSE_lm <- mean(df$SqErr)
MSE_lm 
```
Linear Regression Trendline model has the smallest Mean squared error(MSE). Thus, it is better than other prediction models.


Question 5. 
Calculate a weighted average forecast by averaging out the three forecasts calculated in (3) with the following weights: 4 for trend line, 2 for exponential smoothing, 1 for weighted moving average. Remember to divide by the sum of the weights in a weighted average.
```{r}
df <- c(Ft_wavg, Ft_esp, Ft_lm)
new_w <- c(1, 2, 4)
new_sw <- new_w*df
avg_Forecast <- sum(new_sw)/sum(new_w)
avg_Forecast
```

